Revisiting Child and Adolescent Depression Treatments

### 2024-05-15 Manuscript_for_rendering_in_word.qmd
Final draft

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)

@@

# Introduction

@@

@@

# Method

@@

# Results

@@

df_for_baseline_severity <- df_for_baseline_severity %>% 
  distinct(new_study_id, .keep_all = TRUE) %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

@@

df_for_sex_metan <- df_for_sex_metan %>% 
  filter(percent_women != 100) # filter out studies with only women %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

@@

df_for_age_metan <- df_baseline_demographics %>% 
  filter(!is.na(mean_age)) %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

@@

df_baseline_demographics <- full_join(df_baseline_demographics, df_no_sites, by = "new_study_id") %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

@@

control_summary <- rbind(sessions, frequency, length_sessions_mins, total_hours, total_period_weeks)

@@

control_summary <- control_summary %>%
  mutate(Column = case_when(
    Column == "no_sessions" ~ "Number of sessions",
    Column == "freq_weeks" ~ "Intensity (sessions per week)",
    Column == "length_sessions_mins" ~ "Session length (mins)",
    Column == "total_hours" ~ "Total intervention hours",
    Column == "total_period_weeks" ~ "Intervention duration (weeks)",
    TRUE ~ as.character(Column)  # Keep original value if no match
  ))

control_summary <- control_summary %>% 
  mutate(across(where(is.numeric), ~round(., 2)))

#function for between group effect sizes (cohens d)

effect_btwn <- function(df) {
  pooled_sd <- sqrt(((df$N[1]-1)*df$SD[1]^2 + (df$N[2]-1)*df$SD[2]^2) / (df$N[1] + df$N[2] - 2))
  mean_diff <- df$Mean[1] - df$Mean[2]
  effect_btwn <- mean_diff / pooled_sd
  return(effect_btwn)
}

session_es <- effect_btwn(sessions)
frequency_es <- effect_btwn(frequency)
length_sessions_mins_es <- effect_btwn(length_sessions_mins)
total_hours_es <- effect_btwn(total_hours)
total_period_weeks_es <- effect_btwn(total_period_weeks)

view(df_descr_psy_conditions)
testing <- df_descr_psy_conditions %>% 
  filter()

other_ctr_ids <- df_long_for_metan[df_long_for_metan$type == "other ctr",]$new_study_id
testing <- df_descr_psy_conditions %>% 
  filter(new_study_id %in% other_ctr_ids)
view(testing)
```

as_grouped_data(control_summary, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  bold(i = c(1, 4, 7, 10, 13), bold = TRUE, part = "body") %>%  
  set_table_properties(width = 1) %>%  # Set table width 
  autofit() %>%  # Autofit column widths 
  align(j = 2:4, align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  # add_header_row(values = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", colwidths = 4) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 5", colwidths = 4) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  #italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  #padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all")

@@

session_no_t_test <- t.test(no_sessions ~ levels_var, data = df_descr_psy_conditions)
freq_t_test <- t.test(freq_weeks ~ levels_var, data = df_descr_psy_conditions)
session_length_t_test <- t.test(length_sessions_mins ~ levels_var, data = df_descr_psy_conditions)
total_hours_t_test <- t.test(total_hours ~ levels_var, data = df_descr_psy_conditions)
total_period_t_test <- t.test(total_period_weeks ~ levels_var, data = df_descr_psy_conditions)

@@

results_session_no_t_test <- extract_t_test_stats(session_no_t_test)
results_freq_t_test <- extract_t_test_stats(freq_t_test)
results_session_length_t_test <- extract_t_test_stats(session_length_t_test)
results_total_hours_t_test <- extract_t_test_stats(total_hours_t_test)
results_total_period_t_test <- extract_t_test_stats(total_period_t_test)

control_summary_t_tests <- rbind(results_session_no_t_test, results_freq_t_test, results_session_length_t_test, results_total_hours_t_test, results_total_period_t_test)

# For row names
control_summary_t_tests <- control_summary_t_tests %>%
  mutate(outcome = case_when(
    row_number() == 1 ~ "Number of sessions",
    row_number() == 2 ~ "Intensity (sessions per week)",
    row_number() == 3 ~ "Session length (mins)",
    row_number() == 4 ~ "Total intervention hours",
    row_number() == 5 ~ "Intervention duration (weeks)"
  ))

@@

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
#| label: tbl-t-tests-intensity
#| tbl-cap: "Results for t-tests Comparing the intensity of the intervention between active and control arms of psychotherapy studies"
#| tbl-cap-location: top

flextable(control_summary_t_tests) %>%
  set_table_properties(width = 1) %>%  # Set table width %>% 
  autofit() %>%  # Autofit column widths 
  align(j = 2:ncol(control_summary_t_tests), align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  #add_header_row(values = "Results for t-tests comparing intervention intensity between active and control arms of psychotherapy trials", colwidths = 6) %>% # This is a work around for the problems with rendering captions
  #add_header_row(values = "Table 6", colwidths = 6) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  #italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  #padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all")

@@

As seen in @tbl-intensity-of-int active conditions featured significantly more sessions than control conditions (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Number of sessions"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Number of sessions"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Number of sessions"]`). These sessions were also significantly longer (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Session length (mins)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Session length (mins)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Session length (mins)"]`) and more frequent (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`). The total intervention hours were higher in active vs control conditions (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Total intervention hours"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Total intervention hours"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Total intervention hours"]`), though the duration was similar (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Intervention duration (weeks)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Intervention duration (weeks)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Intervention duration (weeks)"]`).

# Discussion

@@

# Count the occurrences of each type
type_counts <- filtered_df %>%
  count(type) %>%
  mutate(total_count = sum(n),
         wl_cau_sum = sum(n[type %in% c("wl", "cau")]),
         wl_cau_percentage = wl_cau_sum / total_count * 100)

@@

Moving beyond comparison between the two modalities, we examined whether psychotherapy controls are reasonable counterfactuals to receiving the treatment. An obvious disadvantage of psychotherapy trials is that they are typically unblinded (and hard to blind). Yet, our results show that psychotherapy trials are unlikely to fulfill some other basic conditions of the "all else is equal" assumption of a counterfactual condition. In order to test that a psychological treatment is effective per se (e.g. because of the specific techniques) rather than because of generic effects (e.g. pleasant human contact), aspects such as therapist contact time should be matched. We find that there are vast differences between treatment and control arms in psychotherapy trials which may artificially inflate estimates of the efficacy of psychological treatments. Additionally, it is notable that most (67.24%) psychotherapy RCT control conditions are either WL or TAU, both of which are likely to be associated with disappointment bias, thereby further inflating estimates of efficacy of psychological treatments.

@@

Second, the low quality of control conditions of psychotherapy trials for adolescent depression should prompt a rethinking of what needs to be done to create fair comparators. Indeed, investment should be directed into providing rigorous evidence that establishes depression psychotherapies as more efficacious than fair controls. There are examples of psychotherapy RCTs where such rigor has been applied, such as @youngRandomizedDepressionPrevention2016 where active and control arms were matched on frequency and duration of sessions, and total amount of contact.

@@

sessions <- group_function(df_descr_psy_conditions_no_wl,"levels_var","no_sessions")
frequency <- group_function(df_descr_psy_conditions_no_wl,"levels_var","freq_weeks")
length_sessions_mins <- group_function(df_descr_psy_conditions_no_wl,"levels_var","length_sessions_mins")
total_hours <- group_function(df_descr_psy_conditions_no_wl,"levels_var","total_hours")
total_period_weeks <- group_function(df_descr_psy_conditions_no_wl,"levels_var","total_period_weeks")

control_summary_no_wl <- rbind(sessions, frequency, length_sessions_mins, total_hours, total_period_weeks)

@@

control_summary_no_wl <- control_summary_no_wl %>%
  mutate(Column = case_when(
    Column == "no_sessions" ~ "Number of sessions",
    Column == "freq_weeks" ~ "Intensity (sessions per week)",
    Column == "length_sessions_mins" ~ "Session length (mins)",
    Column == "total_hours" ~ "Total intervention hours",
    Column == "total_period_weeks" ~ "Intervention duration (weeks)",
    TRUE ~ as.character(Column)  # Keep original value if no match
  ))

control_summary_no_wl <- control_summary_no_wl %>% 
  mutate(across(where(is.numeric), ~round(., 2)))

session_es_no_wl <- effect_btwn(sessions)
frequency_es_no_wl <- effect_btwn(frequency)
length_sessions_mins_es_no_wl <- effect_btwn(length_sessions_mins)
total_hours_es_no_wl <- effect_btwn(total_hours)
total_period_weeks_es_no_wl <- effect_btwn(total_period_weeks)

as_grouped_data(control_summary_no_wl, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  bold(i = c(1, 4, 7, 10, 13), bold = TRUE, part = "body") %>%  
  set_table_properties(width = 1) %>%  # Set table width 
  autofit() %>%  # Autofit column widths 
  align(j = 2:4, align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  # add_header_row(values = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", colwidths = 4) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 5", colwidths = 4) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  #italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = 1, part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  #padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all")

@@

session_no_t_test <- t.test(no_sessions ~ levels_var, data = df_descr_psy_conditions_no_wl)
freq_t_test <- t.test(freq_weeks ~ levels_var, data = df_descr_psy_conditions_no_wl)
session_length_t_test <- t.test(length_sessions_mins ~ levels_var, data = df_descr_psy_conditions_no_wl)
total_hours_t_test <- t.test(total_hours ~ levels_var, data = df_descr_psy_conditions_no_wl)
total_period_t_test <- t.test(total_period_weeks ~ levels_var, data = df_descr_psy_conditions_no_wl)

@@

results_session_no_t_test <- extract_t_test_stats(session_no_t_test)
results_freq_t_test <- extract_t_test_stats(freq_t_test)
results_session_length_t_test <- extract_t_test_stats(session_length_t_test)
results_total_hours_t_test <- extract_t_test_stats(total_hours_t_test)
results_total_period_t_test <- extract_t_test_stats(total_period_t_test)

control_summary_t_tests_no_wl <- rbind(results_session_no_t_test, results_freq_t_test, results_session_length_t_test, results_total_hours_t_test, results_total_period_t_test)

# For row names
control_summary_t_tests_no_wl <- control_summary_t_tests_no_wl %>%
  mutate(outcome = case_when(
    row_number() == 1 ~ "Number of sessions",
    row_number() == 2 ~ "Intensity (sessions per week)",
    row_number() == 3 ~ "Session length (mins)",
    row_number() == 4 ~ "Total intervention hours",
    row_number() == 5 ~ "Intervention duration (weeks)"
  ))

@@

>>>>

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)
library(effsize)

@@

# Abstract
**Question:** Psychotherapy and medication are the main treatments for adolescent depression, with guidelines recommending psychotherapy as the first-line treatment for most cases. Given the scarcity of head-to-head comparisons, recommendations are based on indirect comparisons of outcomes obtained in randomised controlled trials (RCTs) within each modality. Here we examine whether basic conditions required to draw valid inferences from these indirect comparisons are met.
**Study selection and analysis:** Drawing upon studies included in two recent meta-analyses, and supplementing with an original systematic search, 88 RCTs of psychotherapy and medication for child and adolescent depression (mean age 4-18 years) were identified. Using meta-analytic approaches, we examined comparability of a) participant characteristics and b) trial characteristics in medication and psychotherapy trials. Lastly, we examined whether psychotherapy controls are well-matched to active conditions. 
**Findings:** Participants in medication RCTs had higher depression severity and were more likely to be male compared to psychotherapy RCTs. The pooled effect size of medication controls was very large (SMD=-1.9), and was medium for psychotherapy controls (SMD=-0.5). Medication RCTs also had significantly more trial sites. Within psychotherapy RCTs, control conditions were less intensive on average than active conditions (e.g. fewer contact hours).
**Conclusions:** Medication and psychotherapy RCTs differ on both participant and methodological characteristics, thereby violating the conditions for valid comparison between modalities. Psychotherapy controls often involve little therapist contact and are easy-to-beat comparators. These findings cast doubt on the confidence with which psychotherapy is recommended over medication for youth depression, and highlight the pressing need to improve the evidence base.

# Key messages
Psychotherapy is recommended before medication for most cases of depression in children and adolescents, a recommendation that is based on indirect comparisons of outcomes from randomised controlled trials (RCTs) within each treatment modality. We examine the validity of these inferences by scrutinising the comparability of psychotherapy and medication RCTs. We find significant differences in sample characteristics (namely depression severity and sex composition) and trial design features, such that the effect sizes of medication controls (i.e. pill placebo) are much larger than those for psychotherapy controls, and that medication RCTs feature significantly more trial sites. We also examine the quality of controls used in psychotherapy RCTs and find that they are poorly matched to active intervention arms in ways such as human contact hours, and hence represent poor and easy to beat comparators. Our findings underscore the need for a higher quality evidence base upon which to base treatment guidelines and clinical decision making.

# Background

@@

# Objective

@@

# Study selection and analysis

@@

# Findings

@@

df_for_baseline_severity <- df_for_baseline_severity %>% 
  distinct(new_study_id, .keep_all = TRUE)

df_for_baseline_severity <- df_for_baseline_severity %>% 
    filter(new_study_id != "March, 2004")  # removing TADS because it does not technically belong to either the psy or med groups

@@

df_for_sex_metan <- df_for_sex_metan %>% 
  filter(percent_women != 100) # filter out studies with only women %>%

df_for_sex_metan <- df_for_sex_metan %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

@@

df_for_age_metan <- df_baseline_demographics %>% 
  filter(!is.na(mean_age))

df_for_age_metan <- df_for_age_metan %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

@@

df_baseline_demographics <- full_join(df_baseline_demographics, df_no_sites, by = "new_study_id")

df_baseline_demographics <- df_baseline_demographics %>% 
  filter(new_study_id != "March, 2004") # removing TADS because it does not technically belong to either the psy or med groups

@@

control_summary <- rbind(sessions, frequency, length_sessions_mins, total_hours)

@@

control_summary <- control_summary %>%
  mutate(Column = case_when(
    Column == "no_sessions" ~ "Number of sessions",
    Column == "freq_weeks" ~ "Intensity (sessions per week)",
    Column == "length_sessions_mins" ~ "Session length (mins)",
    Column == "total_hours" ~ "Total intervention hours",
    TRUE ~ as.character(Column)  # Keep original value if no match
  ))

# other_ctr_ids <- df_long_for_metan[df_long_for_metan$type == "other ctr",]$new_study_id
# testing <- df_descr_psy_conditions %>% 
#   filter(new_study_id %in% other_ctr_ids)
# view(testing)

@@

session_no_t_test <- t.test(no_sessions ~ levels_var, data = df_descr_psy_conditions)
freq_t_test <- t.test(freq_weeks ~ levels_var, data = df_descr_psy_conditions)
session_length_t_test <- t.test(length_sessions_mins ~ levels_var, data = df_descr_psy_conditions)
total_hours_t_test <- t.test(total_hours ~ levels_var, data = df_descr_psy_conditions)

@@

results_session_no_t_test <- extract_t_test_stats(session_no_t_test)
results_freq_t_test <- extract_t_test_stats(freq_t_test)
results_session_length_t_test <- extract_t_test_stats(session_length_t_test)
results_total_hours_t_test <- extract_t_test_stats(total_hours_t_test)

# for t test table

control_summary_t_tests <- rbind(results_session_no_t_test, results_freq_t_test, results_session_length_t_test, results_total_hours_t_test)

# For row names
control_summary_t_tests <- control_summary_t_tests %>%
  mutate(outcome = case_when(
    row_number() == 1 ~ "Number of sessions",
    row_number() == 2 ~ "Intensity (sessions per week)",
    row_number() == 3 ~ "Session length (mins)",
    row_number() == 4 ~ "Total intervention hours"
  ))

@@

# function for between group effect sizes (cohens d)
# 
# effect_btwn <- function(df) {
#   pooled_sd <- sqrt(((df$N[1]-1)*df$SD[1]^2 + (df$N[2]-1)*df$SD[2]^2) / (df$N[1] + df$N[2] - 2))
#   mean_diff <- df$Mean[1] - df$Mean[2]
#   effect_btwn <- mean_diff / pooled_sd
#   return(effect_btwn)
# }
# 
# session_es <- effect_btwn(sessions)
# frequency_es <- effect_btwn(frequency)
# length_sessions_mins_es <- effect_btwn(length_sessions_mins)
# total_hours_es <- effect_btwn(total_hours)
# total_period_weeks_es <- effect_btwn(total_period_weeks)

# using a package to calculate instead

session_es <- cohen.d(df_descr_psy_conditions$no_sessions ~ df_descr_psy_conditions$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
frequency_es <- cohen.d(df_descr_psy_conditions$freq_weeks ~ df_descr_psy_conditions$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
length_sessions_mins_es <- cohen.d(df_descr_psy_conditions$length_sessions_mins ~ df_descr_psy_conditions$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )
total_hours_es <- cohen.d(df_descr_psy_conditions$total_hours ~ df_descr_psy_conditions$levels_var, pooled = TRUE, paired = FALSE, within=FALSE, na.rm=FALSE )

# going to try displaying summary statistics and t-tests in one place

control_summary <- cbind(control_summary, 
                          "Cohen's d" = rep(NA, nrow(control_summary)),
                         "Upper CI" = rep(NA, nrow(control_summary)),
                         "Lower CI" = rep(NA, nrow(control_summary)),
                         t = rep(NA, nrow(control_summary)), 
                          df = rep(NA, nrow(control_summary)), 
                          "p-value" = rep(NA, nrow(control_summary)))

# Update the first row with session_es
control_summary[1, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  session_es$estimate, session_es$conf.int[1], session_es$conf.int[2], 
  control_summary_t_tests$`t statistic`[1], control_summary_t_tests$df[1], 
  control_summary_t_tests$`p-value`[1]
)

# Update the third row with frequency_es
control_summary[3, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  frequency_es$estimate, frequency_es$conf.int[1], frequency_es$conf.int[2], 
  control_summary_t_tests$`t statistic`[2], control_summary_t_tests$df[2], 
  control_summary_t_tests$`p-value`[2]
)

# Update the fifth row with length_sessions_mins_es
control_summary[5, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  length_sessions_mins_es$estimate, length_sessions_mins_es$conf.int[1], length_sessions_mins_es$conf.int[2], 
  control_summary_t_tests$`t statistic`[3], control_summary_t_tests$df[3], 
  control_summary_t_tests$`p-value`[3]
)

# Update the seventh row with total_hours_es
control_summary[7, c("Cohen's d", "Upper CI", "Lower CI", "t", "df", "p-value")] <- c(
  total_hours_es$estimate, total_hours_es$conf.int[1], total_hours_es$conf.int[2], 
  control_summary_t_tests$`t statistic`[4], control_summary_t_tests$df[4], 
  control_summary_t_tests$`p-value`[4]
)

control_summary <- control_summary %>%
  mutate(
    `Cohen's d` = as.numeric(`Cohen's d`),
    `Upper CI` = as.numeric(`Upper CI`),
    `Lower CI` = as.numeric(`Lower CI`)
  ) %>%
  mutate(across(where(is.numeric), ~round(., 2)))

# # plot
# 
# ggplot(df_descr_psy_conditions_no_wl, aes(x = total_hours, fill = levels_var)) +
#   geom_histogram(position = "identity", alpha = 0.5, binwidth = 5, color = "black") +
#   facet_wrap(~ levels_var, ncol = 2) +
#   theme_minimal() +
#   labs(title = "Histograms of Number of Total Hours by Group",
#        x = "Number of Hours",
#        y = "Frequency") +
#   scale_fill_manual(values = c("lightpink", "lightblue"))

@@

as_grouped_data(control_summary, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  bold(i = c(1, 4, 7, 10), bold = TRUE, part = "body") %>%  
  set_table_properties(width = 1) %>%  # Set table width 
  autofit() %>%  # Autofit column widths 
  align(j = 2:9, align = "center", part = "all") %>% 
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
  # add_header_row(values = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", colwidths = 4) %>% # This is a work around for the problems with rendering captions
  # add_header_row(values = "Table 5", colwidths = 4) %>% 
  bold(part = "header") %>%   # Bold the relevant header rows
  bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
  #italic(i = 2, italic = TRUE, part = "header") %>% 
  border_remove() %>% # this is to remove borders around the caption that look strange
  hline(j = 1:4, part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
  vline(j = c(1,4), part = "body", border = small_border) %>% 
  border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
  hline(part = "header", border = big_border) %>% 
  #padding(i = 2, padding.bottom = 15, part = "header") %>% 
  paginate(init = TRUE) %>% 
  line_spacing(space = .7, part = "all") 
```

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
# #| label: tbl-t-tests-intensity
# #| tbl-cap: "Results for t-tests Comparing the intensity of the intervention between active and control arms of psychotherapy studies"
# #| tbl-cap-location: top
# 
# flextable(control_summary_t_tests) %>%
#   set_table_properties(width = 1) %>%  # Set table width %>% 
#   autofit() %>%  # Autofit column widths 
#   align(j = 2:ncol(control_summary_t_tests), align = "center", part = "all") %>% 
#   font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri
#   #add_header_row(values = "Results for t-tests comparing intervention intensity between active and control arms of psychotherapy trials", colwidths = 6) %>% # This is a work around for the problems with rendering captions
#   #add_header_row(values = "Table 6", colwidths = 6) %>% 
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
#   #italic(i = 2, italic = TRUE, part = "header") %>% 
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>% 
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>% 
#   #padding(i = 2, padding.bottom = 15, part = "header") %>% 
#   paginate(init = TRUE) %>% 
#   line_spacing(space = .7, part = "all")

@@

<!-- As seen in @tbl-intensity-of-int active conditions featured significantly more sessions than control conditions (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Number of sessions"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Number of sessions"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Number of sessions"]`). These sessions were also significantly longer (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Session length (mins)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Session length (mins)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Session length (mins)"]`) and more frequent (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Intensity (sessions per week)"]`). The total intervention hours were higher in active vs control conditions (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Total intervention hours"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Total intervention hours"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Total intervention hours"]`), though the duration was similar (*t* (`r control_summary_t_tests$df[control_summary_t_tests$Outcome == "Intervention duration (weeks)"]`) = `r control_summary_t_tests$"t statistic"[control_summary_t_tests$Outcome == "Intervention duration (weeks)"]`, *p* = `r control_summary_t_tests$"p-value"[control_summary_t_tests$Outcome == "Intervention duration (weeks)"]`).  -->

As seen in @tbl-intensity-of-int, active conditions featured significantly more sessions when compared to control conditions. Sessions in active conditions were longer and more frequent, resulting in significantly more intervention hours overall. Notably, many control conditions were very poorly described and their intensity could not be quantified, resulting in missing data for a number of trials. We performed a sensitivity analysis where we excluded trials that used a WL control condition; with the exception of number of sessions, differences between active and control arms were no longer statistically significant though remained substantial, with between-group Cohen's ds ranging from 0.41-0.54 (please see @tbl-intensity-of-int-no-wl in the Supplement).

# Conclusions and Clinical Implications

@@

# Count the occurrences of each type
type_counts <- filtered_df %>%
  count(type) %>%
  mutate(total_count = sum(n),
         wl_sum = sum(n[type == "wl"]),
         wl_percentage = wl_sum / total_count * 100)

# how many examples of matching of active and control in number of hours

# instances_count <- df_descr_psy_conditions_no_wl %>%
#   filter(!is.na(total_hours)) %>%
#   group_by(new_study_id, total_hours) %>%
#   summarise(count = n()) %>%
#   filter(count > 1) %>%
#   summarise(total_instances = n())

@@

Moving beyond comparison between the two modalities, we examined whether psychotherapy controls are reasonable counterfactuals to receiving the treatment. An obvious disadvantage of psychotherapy trials is that they are typically unblinded (and hard to blind). Yet, our results show that psychotherapy trials are unlikely to fulfill some other basic conditions of the "all else is equal" assumption of a counterfactual condition. In order to test that a psychological treatment is effective per se (e.g. because of the specific techniques) rather than because of generic effects (e.g. pleasant human contact), aspects such as therapist contact time should be matched. Many (24%) psychotherapy RCTs use waitlist controls, which by definition do not match for hours of therapist contact, and are often associated with disappointment bias. TAU and other psychotherapy control conditions varied drastically; 9 RCTs used controls that exactly matched the active arm in total number of contact hours, though several studies used bibliotherapy (e.g. an educational brochure) or online-only control conditions which did not involve any direct therapist contact. Importantly, controls were often very poorly described, resulting in difficulties quantifying their intensity and hence evaluating their adequacy as counterfactual conditions. Overall, there is a lack of rigor in matching control to active treatment conditions in psychotherapy RCTs, with the latter typically featuring considerably more contact hours, which may artificially inflate estimates of the efficacy of psychological treatments.

@@

Second, the low quality of control conditions of psychotherapy trials for adolescent depression should prompt a rethinking of what needs to be done to create fair comparators. Indeed, investment should be directed into providing rigorous evidence that establishes depression psychotherapies as more efficacious than fair controls. There are examples of psychotherapy RCTs where such rigor has been applied (e.g. Bolton, 2007; Liddle, 1990; Rohde, 2004) in matching active and control arms on variables such as therapist time and attention, provision of homework, and small group interaction.

@@

sessions <- group_function(df_descr_psy_conditions_no_wl,"levels_var","no_sessions")
frequency <- group_function(df_descr_psy_conditions_no_wl,"levels_var","freq_weeks")
length_sessions_mins <- group_function(df_descr_psy_conditions_no_wl,"levels_var","length_sessions_mins")
total_hours <- group_function(df_descr_psy_conditions_no_wl,"levels_var","total_hours")

control_summary_no_wl <- rbind(sessions, frequency, length_sessions_mins, total_hours)

@@

control_summary_no_wl <- control_summary_no_wl %>%
  mutate(Column = case_when(
    Column == "no_sessions" ~ "Number of sessions",
    Column == "freq_weeks" ~ "Intensity (sessions per week)",
    Column == "length_sessions_mins" ~ "Session length (mins)",
    Column == "total_hours" ~ "Total intervention hours",
    TRUE ~ as.character(Column)  # Keep original value if no match
  ))

@@

session_no_t_test <- t.test(no_sessions ~ levels_var, data = df_descr_psy_conditions_no_wl)
freq_t_test <- t.test(freq_weeks ~ levels_var, data = df_descr_psy_conditions_no_wl)
session_length_t_test <- t.test(length_sessions_mins ~ levels_var, data = df_descr_psy_conditions_no_wl)
total_hours_t_test <- t.test(total_hours ~ levels_var, data = df_descr_psy_conditions_no_wl)

@@

results_session_no_t_test <- extract_t_test_stats(session_no_t_test)
results_freq_t_test <- extract_t_test_stats(freq_t_test)
results_session_length_t_test <- extract_t_test_stats(session_length_t_test)
results_total_hours_t_test <- extract_t_test_stats(total_hours_t_test)

control_summary_t_tests_no_wl <- rbind(results_session_no_t_test, results_freq_t_test, results_session_length_t_test, results_total_hours_t_test)

# For row names
control_summary_t_tests_no_wl <- control_summary_t_tests_no_wl %>%
  mutate(outcome = case_when(
    row_number() == 1 ~ "Number of sessions",
    row_number() == 2 ~ "Intensity (sessions per week)",
    row_number() == 3 ~ "Session length (mins)",
    row_number() == 4 ~ "Total intervention hours"
  ))

@@

# calculating with my function

# session_es_no_wl <- effect_btwn(sessions)
# frequency_es_no_wl <- effect_btwn(frequency)
# length_sessions_mins_es_no_wl <- effect_btwn(length_sessions_mins)
# total_hours_es_no_wl <- effect_btwn(total_hours)
# total_period_weeks_es_no_wl <- effect_btwn(total_period_weeks)

# using a package to calculate instead
